Robinson et al. 2020 Supplemental Information: Comparing the Radiocarbon and Tree-ring Records for the Upland US Southwest

### 2020-01-19 analysis/paper/paper.Rmd
reduced memory overhead

```{r prepare-radiocarbon-data}
# # Read in the radiocarbon database
# # radiocarbon_data <-
# readr::read_csv(
#   here::here("analysis/data/raw_data/0403_sbox_clean.csv")
# ) %>%
#   # Convert to spatial object
#   sf::st_as_sf(coords = c("Long","Lat"),
#                crs = 4326) %>%
#   sf::st_transform(four_corners_lcc_proj) %>%
#   # Select salient columns
#   dplyr::select(labnumber,
#                 date,
#                 sd,
#                 Site,
#                 SiteID) %>%
#   dplyr::rename(Lab_Number = labnumber,
#                 Date_BP = date,
#                 SD = sd,
#                 Site_Number = Site) %>%
#   sf::st_intersection(uusw_counties) %>%
#   sf::st_drop_geometry() %>%
#   dplyr::mutate(Type = "Radiocarbon",
#                 SiteID = SiteID %>% stringr::str_c("Radiocarbon - ", .)) %>%
#   dplyr::select(Type,
#                 SiteID,
#                 Site_Number,
#                 County,
#                 Lab_Number,
#                 Date_BP,
#                 SD) %T>%
#   # Write the output to a new file (csv)
#   readr::write_csv(
#     here::here("analysis/data/derived_data/radiocarbon_data.csv")
#   )

@@

```{r prepare-treering-data}
# # Read in the Bocinsky et al. 2016 tree ring data
# # https://doi.org/10.6067/XCV82J6D7B
# dendro_data <-
#   readr::read_csv(
#     here::here("analysis/data/raw_data/BOCINSKY_ET_AL_2015_TREE_RINGS.csv")
#   ) %>%
#   # Convert to spatial object
#   sf::st_as_sf(coords = c("LON","LAT"),
#                crs = 4326) %>%
#   sf::st_transform(four_corners_lcc_proj) %>%
#   # Convert dates to BP
#   dplyr::mutate(Date_BP = 1950 - Outer_Date_AD,
#                 Type = ifelse(C_Level %in% c(2,3),
#                               "Cutting or Near-cutting",
#                               "Non-cutting")) %>%
#   # drop C_Level
#   dplyr::select(Seq_Number,
#                 Site_Number,
#                 Lab_Number,
#                 Type,
#                 Date_BP) %>%
#   dplyr::rename(SiteID = Seq_Number) %>%
#   sf::st_intersection(uusw_counties) %>%
#   sf::st_drop_geometry() %>%
#   dplyr::mutate(Type = Type %>% stringr::str_c("Dendro - ", .),
#                 SiteID = SiteID %>% stringr::str_c("Dendro - ", .),
#                 SD = 0) %>%
#   dplyr::filter(Type == "Dendro - Cutting or Near-cutting") %>%
#   dplyr::mutate(Type = "Tree-ring")  %>%
#   dplyr::select(Type,
#                 SiteID,
#                 Site_Number,
#                 County,
#                 Lab_Number,
#                 Date_BP,
#                 SD) %T>%
#   # Write the output to a new file (csv)
#   readr::write_csv(
#     here::here("analysis/data/derived_data/dendro_data.csv")
#   )

@@

```{r radiocarbon-spds}
# Set the time range for the reconstructions
timerange <- c(1949,0)

if(!file.exists(
  here::here("analysis/data/derived_data/radiocarbon_spd.rds")
))
  all_data$Radiocarbon %>%
  split(.,.$`Study Area`) %>%
  tibble::tibble(`Study Area` = names(.),
                 Dates = .) %>%
  dplyr::bind_rows(tibble::tibble(`Study Area` = "UUSW",
                                  Dates = list(all_data$Radiocarbon))) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(Calibrated = list(Dates %$%
                                    rcarbon::calibrate(x = Date_BP,
                                                       errors = SD,
                                                       timerange = timerange,
                                                       calMatrix = TRUE,
                                                       verbose = FALSE)),
                Bins = list(rcarbon::binPrep(sites = Dates$SiteID,
                                             ages = Calibrated,
                                             h = 100)),
                SPD = list(rcarbon::spd(x = Calibrated,
                                        bins = Bins,
                                        datenormalised = TRUE,
                                        spdnormalised = TRUE,
                                        timeRange = c(1949,0),
                                        runm = 21,
                                        verbose = FALSE))) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(Density = SPD %>%
                  purrr::map(function(x){
                    x$grid %>%
                      tibble::as_tibble() %$%
                      tibble::tibble(Type = "Radiocarbon Density",
                                     Date_BP = calBP,
                                     Density = PrDens / sum(PrDens, na.rm = T))
                  })
  ) %>%
  readr::write_rds(
    here::here("analysis/data/derived_data/radiocarbon_spd.rds"),
    compress = "gz")

@@

```{r treering-spds}
calibrate_tree_ring_dates <- function(x){
  out <- list(
    metadata = tibble::tibble(
      DateID = 1:length(x),
      CRA = x,
      Error = 0,
      Details = NA,
      CalCurve = "intcal13",
      ResOffsets = 0,
      ResErrors = 0,
      StartBP = timerange[1],
      EndBP = 0,
      Normalised = TRUE,
      F14C = FALSE,
      CalEPS = 1e-05
    ),
    grids = NA,
    calmatrix = 
      tibble::tibble(index = 1:length(x),
                     year_bp = x,
                     value = TRUE) %>%
      tidyr::pivot_wider(names_from = index,
                         values_from = value) %>%
      dplyr::bind_rows(tibble::tibble(year_bp = timerange[1]:timerange[2])) %>%
      dplyr::filter(!duplicated(year_bp)) %>%
      dplyr::arrange(-year_bp) %>%
      dplyr::filter(year_bp >= 0) %>%
      tibble::column_to_rownames("year_bp") %>%
      as.matrix() %>%
      tidyr::replace_na(0) %>%
      magrittr::set_colnames(NULL)
  )

@@

if(!file.exists(
  here::here("analysis/data/derived_data/dendro_spd.rds")
))
  all_data$`Tree-ring` %>%
  split(.,.$`Study Area`) %>%
  tibble::tibble(`Study Area` = names(.),
                 Dates = .) %>%
  dplyr::bind_rows(tibble::tibble(`Study Area` = "UUSW",
                                  Dates = list(all_data$`Tree-ring`))) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(Calibrated = list(Dates %$%
                                    calibrate_tree_ring_dates(x = Date_BP)),
                Bins = list(rcarbon::binPrep(sites = Dates$SiteID,
                                             ages = Dates$Date_BP,
                                             h = 100)),
                SPD = list(rcarbon::spd(x = Calibrated,
                                        bins = Bins,
                                        datenormalised = TRUE,
                                        spdnormalised = TRUE,
                                        timeRange = timerange,
                                        runm = 21,
                                        verbose = FALSE))) %>%

>>>>

```{r prepare-radiocarbon-data}
# Read in the radiocarbon database
if(!file.exists(
  here::here("analysis/data/derived_data/radiocarbon_data.csv")
))
  readr::read_csv(
    here::here("analysis/data/raw_data/0403_sbox_clean_sd300.csv")
  ) %>%
  # Convert to spatial object
  sf::st_as_sf(coords = c("Long","Lat"),
               crs = 4326) %>%
  sf::st_transform(four_corners_lcc_proj) %>%
  # Select salient columns
  dplyr::select(labnumber,
                date,
                sd,
                Site,
                SiteID) %>%
  dplyr::rename(Lab_Number = labnumber,
                Date_BP = date,
                SD = sd,
                Site_Number = Site) %>%
  sf::st_intersection(uusw_counties) %>%
  sf::st_drop_geometry() %>%
  dplyr::mutate(Type = "Radiocarbon",
                SiteID = SiteID %>% stringr::str_c("Radiocarbon - ", .)) %>%
  dplyr::select(Type,
                SiteID,
                Site_Number,
                County,
                Lab_Number,
                Date_BP,
                SD) %>%
  # Write the output to a new file (csv)
  readr::write_csv(
    here::here("analysis/data/derived_data/radiocarbon_data.csv")
  )

@@

```{r prepare-treering-data}
# Read in the Bocinsky et al. 2016 tree ring data
# https://doi.org/10.6067/XCV82J6D7B

if(!file.exists(
  here::here("analysis/data/derived_data/dendro_data.csv")
))
  readr::read_csv(
    here::here("analysis/data/raw_data/BOCINSKY_ET_AL_2015_TREE_RINGS.csv")
  ) %>%
  # Convert to spatial object
  sf::st_as_sf(coords = c("LON","LAT"),
               crs = 4326) %>%
  sf::st_transform(four_corners_lcc_proj) %>%
  # Convert dates to BP
  dplyr::mutate(Date_BP = 1950 - Outer_Date_AD,
                Type = ifelse(C_Level %in% c(2,3),
                              "Cutting or Near-cutting",
                              "Non-cutting")) %>%
  # drop C_Level
  dplyr::select(Seq_Number,
                Site_Number,
                Lab_Number,
                Type,
                Date_BP) %>%
  dplyr::rename(SiteID = Seq_Number) %>%
  sf::st_intersection(uusw_counties) %>%
  sf::st_drop_geometry() %>%
  dplyr::mutate(Type = Type %>% stringr::str_c("Dendro - ", .),
                SiteID = SiteID %>% stringr::str_c("Dendro - ", .),
                SD = 0) %>%
  dplyr::filter(Type == "Dendro - Cutting or Near-cutting") %>%
  dplyr::mutate(Type = "Tree-ring")  %>%
  dplyr::select(Type,
                SiteID,
                Site_Number,
                County,
                Lab_Number,
                Date_BP,
                SD) %>%
  # Write the output to a new file (csv)
  readr::write_csv(
    here::here("analysis/data/derived_data/dendro_data.csv")
  )

@@

```{r radiocarbon-spds}
# Set the time range for the reconstructions
timerange <- c(1949, 0)

all_data$Radiocarbon %>%
  split(.,.$`Study Area`) %>%
  tibble::tibble(`Study Area` = names(.),
                 Dates = .) %>%
  dplyr::bind_rows(tibble::tibble(`Study Area` = "UUSW",
                                  Dates = list(all_data$Radiocarbon))) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(Calibrated = list(Dates %$%
                                    rcarbon::calibrate(x = Date_BP,
                                                       errors = SD,
                                                       timeRange = timerange,
                                                       calMatrix = TRUE,
                                                       verbose = FALSE)),
                Bins = list(rcarbon::binPrep(sites = Dates$SiteID,
                                             ages = Calibrated,
                                             h = 100)),
                SPD = list(rcarbon::spd(x = Calibrated,
                                        bins = Bins,
                                        spdnormalised = TRUE,
                                        timeRange = timerange,
                                        runm = 21,
                                        verbose = FALSE))) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(Density = SPD %>%
                  purrr::map(function(x){
                    x$grid %>%
                      tibble::as_tibble() %$%
                      tibble::tibble(Type = "Radiocarbon Density",
                                     Date_BP = calBP,
                                     Density = PrDens / sum(PrDens, na.rm = T))
                  })
  ) %>%
  readr::write_rds(
    here::here("analysis/data/derived_data/radiocarbon_spd.rds"),
    compress = "gz")

@@

```{r treering-spds}
calibrate_tree_ring_dates <- function(x){
  out <- list(
    metadata = tibble::tibble(
      DateID = 1:length(x),
      CRA = x,
      Error = 0,
      Details = NA,
      CalCurve = "intcal13",
      ResOffsets = 0,
      ResErrors = 0,
      StartBP = timerange[1],
      EndBP = timerange[2],
      Normalised = TRUE,
      F14C = FALSE,
      CalEPS = 1e-05
    ),
    grids = NA,
    calmatrix = 
      tibble::tibble(index = 1:length(x),
                     year_bp = x,
                     value = TRUE) %>%
      tidyr::pivot_wider(names_from = index,
                         values_from = value) %>%
      dplyr::bind_rows(tibble::tibble(year_bp = timerange[1]:timerange[2])) %>%
      dplyr::filter(!duplicated(year_bp)) %>%
      dplyr::arrange(-year_bp) %>%
      dplyr::filter(year_bp >= 0) %>%
      tibble::column_to_rownames("year_bp") %>%
      as.matrix() %>%
      tidyr::replace_na(0) %>%
      magrittr::set_colnames(NULL)
  )

@@

all_data$`Tree-ring` %>%
  split(.,.$`Study Area`) %>%
  tibble::tibble(`Study Area` = names(.),
                 Dates = .) %>%
  dplyr::bind_rows(tibble::tibble(`Study Area` = "UUSW",
                                  Dates = list(all_data$`Tree-ring`))) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(Calibrated = list(Dates %$%
                                    calibrate_tree_ring_dates(x = Date_BP)),
                Bins = list(rcarbon::binPrep(sites = Dates$SiteID,
                                             ages = Dates$Date_BP,
                                             h = 100)),
                SPD = list(rcarbon::spd(x = Calibrated,
                                        bins = Bins,
                                        spdnormalised = TRUE,
                                        timeRange = timerange,
                                        runm = 21,
                                        verbose = FALSE))) %>%
